Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2018-01-24 papers/integration/omm_integ.tex
A couple of further corrections based on the final proofing process

% Try to make this clearer for Reviewer 2, Methods Comments, point 5.
\rvwt{To understand the need for this invariance,
consider the effect of a 2D Gaussian hill in SC\_deep which elicits a
successful horizontal saccade of 10\dg~using the weight maps shown in
Figures 7(a)--(c). Activity from the 2D Gaussian (schematically
represented as the large purple dashed circle in Figure 7C),
passing through the weight maps will excite the superior (not shown) and inferior
rectus [Figure 7(b)] channels by an equal, balanced amount, so these cancel out,
allowing the eye movement to be horizontal. The amount of activation
passed to the lateral rectus muscle results from a convolution of the
Gaussian and the exponential component of the weight map relationship
in Eq.~\ref{eq:tabareau_eq3}. If the Gaussian hill now appears further
along the collicular surface, coding for a 20\dg~saccade, \e{and also
becomes smaller} (small purple dashed circle), we can still argue that the vertical component
signals to superior/inferior rectus muscles will cancel out, and we
could imagine that the exponential component of
Eq.~\ref{eq:tabareau_eq3} is correctly parameterised to compensate for
the smaller hill. Now consider a 2-D Gaussian hill which
codes for a 10\dg~saccade which is `down, and to the left' in equal
proportions (large red circle). That means that the hill will sit on the boundary between
the weight maps for the `down', and the `left' muscles. Now, if the
hill moves to the $r=\mathrm{20}$\dg~location on the colliculus (small
red circle), \e{and also
reduces in size}, it will excite only the periphery of the sine; the
exponential increase of the map along $r$ is not guaranteed to
compensate for the reduction in the convolution of the Gaussian hill
and the sinusoidally varying component of the weight map along the $\phi$ axis in
Figure 7(a).}
% Add figure for this.

This led us to hypothesise that the retinotopic mapping \rvwt{to the
SBG be preceded} by an associated widening projection field such that
the hill of activity in \rvwt{a `final' deep layer of superior
colliculus is invariant with position on the map}. There are a number of
locations in the system in which this widening projection field could
exist. It could be implemented in the projections between the retinal
populations and the superficial layer of SC along with the projection
between the World and the FEF population. However, this would affect
activity within the basal ganglia of the model, contradicting a result
in \cite{cope_basal_2017} which explains the `hockey stick' profile
for saccade latency as a function of saccade eccentricity. Instead, we
suggest that a widening projection field is encoded within the
superior colliculus itself, a complex, multi-layered structure which
could quite plausibly support such a function. Indeed, such widening
activity can be seen in the stimulation experiments
in \cite{vokoun_intralaminar_2010} and \cite{vokoun_response_2014}.
%
\rvw{In \cite{ghitani_excitatory_2014}, from the same research group,
evidence is presented for an excitatory and widely projecting pathway
from the stratum griseum intermediale (equivalent to our SC\_deep) to
the more superficial layers stratum opticum and stratum griseum
superficiale. Although this pathway is a `wide' projecting field, the
experiments do not indicate whether the projection \e{widens} along
the rostral-caudal axis of the SC. \cite{bayguinov_hard-wired_2015}
present evidence for another projecting field within SC whose
connectivity pattern \e{does} change along the rostral-caudal
axis. This projection is inhibitory in nature. Although neither of
these results precisely match the widening, excitatory projection
field hypothesised here, they do indicate that such connection
patterns are plausible.}
%
Although in this work we do not model
the SC in detail, we extended the model with a third functional layer
named SC\_deep2, shown in Fig.~\ref{fig:scdeep}(b) (\ccg~has only the two
layers SC\_sup and SC\_deep). We introduced a widening projection
based on a Gaussian projection field whose width, $\sigma(r)$ varies
in inverse proportion to the magnification factor, $M(r)$, given in
Eq.~\ref{eq:cmf} according to:
\begin{equation} \label{eq:sigmar}
\sigma(r) = \frac{m_{\sigma}}{M(r)} - \frac{m_{\sigma}}{M^0} + \sigma_0 \qquad r > r_0
\end{equation}
$m_{\sigma}$ is a scalar parameter which determines the `magnitude of
the widening'. $M^0$ is the `starting' magnification factor. Within
the foveal region ($0 \leq r \leq r_0$), the projection field is not
allowed to widen and so
\begin{equation} \label{eq:sigmar2}
\sigma(r) = \sigma_0 \qquad r \leq r_0
\end{equation}
which makes $\sigma_0$ the width of the Gaussian projection field
within the foveal region. (Note that the value chosen for the width of
the foveal region, $r_0$ is not identical to the foveal shift
parameter used in the \e{DecayingAtFovea} projections into striatum.)
The \e{Widening Gaussian} projection weight, $w(r,d)$ is then computed
as:
\begin{equation} \label{eq:widening}
w(r,d) = e^{-\frac{d^2}{2\sigma\left(r\right)^2}}
\end{equation}
where $d$ is the distance between the source and destination elements
in the collicular plane.  $m_\sigma$ was set to 50, $\sigma_0$ was
0.3, $M^0$ was 12.43 and $r_0$ was 20.
% Could have a figure here, made from widening_gaussian.m

>>>>

% Try to make this clearer for Reviewer 2, Methods Comments, point 5.
\rvwt{To understand the need for this invariance,
consider the effect of a 2D Gaussian hill in SC\_deep which elicits a
successful horizontal saccade of 10\dg~using the weight maps shown in
Figs.~\ref{fig:scdeep}(a)--(c). Activity from the 2D Gaussian
(schematically represented as the large purple dashed circle in
Fig.~\ref{fig:scdeep}(c)), passing through the weight maps will excite
the superior (not shown) and inferior rectus
[Fig.~\ref{fig:scdeep}(b)] channels by an equal, balanced amount, so
these cancel out, allowing the eye movement to be horizontal. The
amount of activation passed to the lateral rectus muscle results from
a convolution of the Gaussian and the exponential component of the
weight map relationship in Eq.~\ref{eq:tabareau_eq3}. If the Gaussian
hill now appears further along the collicular surface, coding for a
20\dg~saccade, \e{and also becomes smaller} (small purple dashed
circle), we can still argue that the vertical component signals to
superior/inferior rectus muscles will cancel out, and we could imagine
that the exponential component of Eq.~\ref{eq:tabareau_eq3} is
correctly parameterised to compensate for the smaller hill. Now
consider a 2-D Gaussian hill which codes for a 10\dg~saccade which is
`down, and to the left' in equal proportions (large red circle). That
means that the hill will sit on the boundary between the weight maps
for the `down', and the `left' muscles. Now, if the hill moves to the
$r=\mathrm{20}$\dg~location on the colliculus (small red
circle), \e{and also reduces in size}, it will excite only the
periphery of the sine; the exponential increase of the map along $r$
is not guaranteed to compensate for the reduction in the convolution
of the Gaussian hill and the sinusoidally varying component of the
weight map along the $\phi$ axis in Fig.~\ref{fig:scdeep}(a).}
% Add figure for this.

This led us to hypothesise that the retinotopic mapping \rvwt{to the
SBG be preceded} by an associated widening projection field such that
the hill of activity in \rvwt{a `final' deep layer of superior
colliculus is invariant with position on the map}. There are a number of
locations in the system in which this widening projection field could
exist. It could be implemented in the projections between the retinal
populations and the superficial layer of SC along with the projection
between the World and the FEF population. However, this would affect
activity within the basal ganglia of the model, contradicting a result
in \cite{cope_basal_2017} which explains the `hockey stick' profile
for saccade latency as a function of saccade eccentricity. Instead, we
suggest that a widening projection field is encoded within the
superior colliculus itself, a complex, multi-layered structure which
could quite plausibly support such a function. Indeed, such widening
activity can be seen in the stimulation experiments
in \cite{vokoun_intralaminar_2010} and \cite{vokoun_response_2014}.
%
\rvw{In \cite{ghitani_excitatory_2014}, from the same research group,
evidence is presented for an excitatory and widely projecting pathway
from the stratum griseum intermediale (equivalent to our SC\_deep) to
the more superficial layers stratum opticum and stratum griseum
superficiale. Although this pathway is a `wide' projecting field, the
experiments do not indicate whether the projection \e{widens} along
the rostral-caudal axis of the SC. \cite{bayguinov_hard-wired_2015}
present evidence for another projecting field within SC whose
connectivity pattern \e{does} change along the rostral-caudal
axis. This projection is inhibitory in nature. Although neither of
these results precisely match the widening, excitatory projection
field hypothesised here, they do indicate that such connection
patterns are plausible.}
%
Although in this work we do not model
the SC in detail, we extended the model with a third functional layer
named SC\_deep2, shown in Fig.~\ref{fig:scdeep}(d) (\ccg~has only the two
layers SC\_sup and SC\_deep). We introduced a widening projection
based on a Gaussian projection field whose width, $\sigma(r)$ varies
in inverse proportion to the magnification factor, $M(r)$, given in
Eq.~\ref{eq:cmf} according to:
\begin{equation} \label{eq:sigmar}
\sigma(r) = \frac{m_{\sigma}}{M(r)} - \frac{m_{\sigma}}{M^0} + \sigma_0 \qquad r > r_0
\end{equation}
$m_{\sigma}$ is a scalar parameter which determines the `magnitude of
the widening'. $M^0$ is the `starting' magnification factor. Within
the foveal region ($0 \leq r \leq r_0$), the projection field is not
allowed to widen and so
\begin{equation} \label{eq:sigmar2}
\sigma(r) = \sigma_0 \qquad r \leq r_0
\end{equation}
which makes $\sigma_0$ the width of the Gaussian projection field
within the foveal region. (Note that the value chosen for the width of
the foveal region, $r_0$ is not identical to the foveal shift
parameter used in the \e{DecayingAtFovea} projections into striatum.)
The \e{Widening Gaussian} projection weight, $w(r,d)$ is then computed
as:
\begin{equation} \label{eq:widening}
w(r,d) = e^{-\frac{d^2}{2\sigma\left(r\right)^2}}
\end{equation}
where $d$ is the distance between the source and destination elements
in the collicular plane.  $m_\sigma$ was set to 50, $\sigma_0$ was
0.3, $M^0$ was 12.43 and $r_0$ was 20.
% Could have a figure here, made from widening_gaussian.m
